\textbf{\sffamily\Huge \ShortTitle}\\ {\textbf{\sffamily\Large \FullTitle}} \vspace{1cm}

### 2012-12-29 paper/mswp2012-proposal.tex
Most ethical project ever

\note{
\begin{itemize}
\item {\em Length: 1-2 pages}
\item Provide a brief explanation of the ethical issue involved and how it will be dealt with appropriately.
\item Are there any security-sensitive issues that apply to your proposal?
\end{itemize}
}

>>>>

%\note{
%\begin{itemize}
%\item {\em Length: 1-2 pages}
%\item Provide a brief explanation of the ethical issue involved and how it will be dealt with appropriately.
%\item Are there any security-sensitive issues that apply to your proposal?
%\end{itemize}
%}

Whenever personal data is stored in a centralized system discussions about security and privacy arise - this is especially true for sensible data like illnesses and health records. Since the modern health care system deals with very sensible and personal data, security and privacy are two very important issue to deal with and we need to take them seriously. \\

By combining data from various different sources the modern health care system is able to create a detailed health report of each and every participant and we need to make sure that these reports are anonymized and secured against attacks from outside as well as inside. From an ethical point of view the modern health care system performs research on humans and the collected data threatens the participants' privacy. \\

The modern health care system performs research on humans and involves patients and human data collection. Therefore ethical issues arise. Ethics aren't to be confused with behaving in accordance with social conventions, beliefs and laws - Tomas Paul and Linda Elder define ethics as \emph{a set of concepts and principles that guide us in determining what behavior helps or harms sentient creatures} [\cite{paul2006thinker}]. \\

The ethical issues that arise for the modern health care system are the following:

\begin{itemize}
	\item \textbf{Respect for autonomy}: the participant has the right to stop the participation whenever he wants to. All data associated with his record needs to be permanetly deleted and removed from our databases.
	\item \textbf{Respect for persons}: the participant has the right to be treated with dignity.
	\item \textbf{Right for inquiry}: the participant has the right to be informed about all the data associated with his record.
	\item \textbf{Truthfulness and honesty}: the participant must not be misinformed about the information derived from his record and how this information was derived.
	\item \textbf{Privacy} the participant has the right that all the information associated with his record is kept private and secured against unpurposeful usage.
\end{itemize}

We deal with these issues accordingly:

\paragraph{Respect for autonomy}
The modern health care system provides a way for each participant to exit our study and remove all his data from our records by the means of an online form. The participant can optionally attach information about the reasons why he is leaving the study.

\paragraph{Respect for persons}
Every participant of the modern health care system is treated in the same way, regardless of his health condition, culture, religion or any other data associated with his record.

\paragraph{Right for inquiry}
The modern health care system provides an online form similar to current social networks that enables the participants to receive a detailed record about all the data associated with their record.

\paragraph{Truthfulness and honesty}
The record containing the information about all the data associated with a participants record is created as detailed as possible. Furthermore, whenever the modern health care system derives a possible health risk from the participants condition and actions recorded an information is provided why the system considers the current situation a risk.

\paragraph{Privacy}
Data privacy is ensured by today's means. We use state-of-the-art security and privacy algorithms to keep our data as secure as possible. Of course, as always, security can't be fully guaranteed, but we highly respect our responsibility of keeping all the data as secure as possible. By picking the most suitable algoriths for data obsfuscation like Partitioning, Randomization or Group Based Anonymization we make it as hard as possible to associate given data with a specific participant. Furthermore all our data is stored encrypted and the communication over the internet is performed encrypted as well as over secure channels.
